\mytitle{}

### 2018-11-12 paper/paper.tex
Some fixes according to Simon

\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	pdftitle={\mytitle{}},
%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros}
	pdfauthor={Anonymous Giraffe and Badger}
}
\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }

@@

\title{\mytitle{}}
\author{
%	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
%	Dimitrios P.\ Pezaros\\
%	University of Glasgow, Glasgow, Scotland,\\
%	\email{k.simpson.1@research.gla.ac.uk}
Anonymous Giraffe,
Anonymous Badger\\
Unnamed Department, Nowhere\\
\email{giraffe.a@unnamed.com}
}

@@

While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems---to quite a large extent, this remains the case today \cite{DBLP:conf/sp/SommerP10}.
Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
Briefly, these are the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
Further issues arise from the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges encountered when learning from unlabelled (often partial) data.
All of these factors greatly inflate the difficulty of the detection problem.

@@

In service of this goal, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling for deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention, improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

This combination of algorithm and coding strategy is well-optimised.
If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
Action values for any state are then obtained by summation of the weight vectors from all activated tiles.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
This means that we need only perform $n_{\mathit{tilings}}$ floating point additions per model update:
\begin{equation}
	\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
	\label{eqn:sg-sarsa-opt}
\end{equation}
If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.

@@

\begin{figure}
	\includegraphics[width=0.8\linewidth]{../plots/ftprep-good-2}
	\caption{
		Demonstration of learning rate for several features included as part of \cref{fig:udp-feature-plots}.
		Performance starts low, before achieving convergence at some later time; this explains the high volume of outliers occurring below the main distribution of many features.
		\label{fig:udp-fp-converge}
	}
\end{figure}

@@

\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point averaged over 10 episodes.
%?? Core findings---different protocols need different features, so everything we proposed above has a use!
The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
%?? How do they do when combined after individual training? Pretty well, especially for TCP.
Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.

@@

%?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for.
We suspect that some of these difficulties originate from the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to unpick the true value of each action), and controlling shared global state.
It may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows, to explore this hypothesis.

@@

%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are an improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
The benefits we have at present are, however, substantial.
What we offer above the alternate approaches we discussed in \cref{sec:related-work} are potentially more flexible deployments and the ability to adapt by design to changing network traffic characteristics, without requiring the network resources or capabilities that the most effective techniques rely upon.

@@

Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies---\SIrange{100}{120}{\kibi\byte} from our experiments.
Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
We believe that it may be possible to implement similar behaviour through application of programmable data planes.

@@

While this work still trails behind the performance of exact DDoS flow detection mechanisms, we hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
Yet even now, we offer advantages above these works, such as online adaptation and potentially more flexible deployment, without requiring many of the same network resources or capabilities that they rely upon.
It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.
%(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?

@@

%?? IDEA: try out average reward, TD($\lambda$) methods as future work...
A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation.

@@

\section*{Acknowledgements}
%Our thanks go to Simon Rogers, Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Richard Cziva for their advice, comments and technical assistance.
Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
%Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
This work was supported by a nameless funder [grant number 0].

>>>>

\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	pdftitle={\mytitle{}},
%	pdfauthor={Kyle A. Simpson, Dimitrios P. Pezaros, Simon D. Rogers}
	pdfauthor={Anonymous Giraffe, Anonymous Badger, Anonymous Ibex}
}
\newcommand*{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}} }

@@

\title{\mytitle{}}
\author{
%	Kyle A.\ Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
%	Kyle A.\ Simpson,
%	Dimitrios P.\ Pezaros,
%   Simon D.\ Rogers\\
%	University of Glasgow, Glasgow, Scotland,\\
%	\email{k.simpson.1@research.gla.ac.uk}
Anonymous Giraffe,
Anonymous Badger,
Anonymous Ibex\\
Unnamed Department, Nowhere\\
\email{giraffe.a@unnamed.com}
}

@@

While \emph{machine learning} (ML) approaches seem like a sensible fit for this problem, in \citeyear{DBLP:conf/sp/SommerP10} \citeauthor{DBLP:conf/sp/SommerP10} identified the `failure to launch' of ML-based anomaly detection systems, and to quite a large extent this remains the case today \cite{DBLP:conf/sp/SommerP10}.
Their application is made difficult due to significant operational differences from standard ML tasks, alongside certain characteristics of network traffic.
Briefly, these are the diversity of network traffic across varying timescales \cite{DBLP:conf/sp/SommerP10} and significant burstiness \cite{DBLP:journals/ccr/LelandWTW95}.
Above the aggregate level, the constant deployment of new services and new protocols means that traffic is \emph{non-stationary} and displays an evolving notion of normality.
Further issues arise from the extraordinarily low tolerance for false positives inherent to network intrusion detection \cite{DBLP:conf/ccs/Axelsson99}, and the challenges encountered when learning from unlabelled (often partial) data.
All of these factors greatly inflate the difficulty of the detection problem.

@@

In service of this goal, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling for deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention, improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

This combination of algorithm and coding strategy is well-optimised.
If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}})$ floating point additions per decision.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
This means that we need only perform $n_{\mathit{tilings}} + 2$ floating point additions and \num{2} multiplications per model update:
\begin{equation}
	\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
	\label{eqn:sg-sarsa-opt}
\end{equation}
If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{../plots/ftprep-good-2}
	\caption{
		Demonstration of learning rate for several features included as part of \cref{fig:udp-feature-plots}.
		Performance starts low, before achieving convergence at some later time; this explains the high volume of outliers occurring below the main distribution of many features.
		\label{fig:udp-fp-converge}
	}
\end{figure}

@@

\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point averaged over 10 episodes.
%?? Core findings---different protocols need different features, so everything we proposed above has a use!
The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
All features converge to their highest-observed performance within around \num{4000} timesteps: \cref{fig:udp-fp-converge} shows an example of this behaviour for differently performing features.
In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
%?? How do they do when combined after individual training? Pretty well, especially for TCP.
Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.

@@

%?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for.
%?? I don’t know if you want to also mention the problem of things getting “stuck” in locally optimal (but globally sub-optimal) policies. I think it’s fairly uncontroversial that an increase in state space would make this more likely. Also, the lack of discounting could play a large role (and more so as the state space becomes more complex…
The most likely explanation is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
The increased state-space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state-space.
We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to unpick the true value of each action), and controlling shared global state.
It may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows, to explore this hypothesis.

@@

%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are an improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
The benefits we have at present are, however, substantial.
What we offer above the alternate approaches we discussed in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.

@@

Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies---\SIrange{100}{120}{\kibi\byte} from our experiments.
As we have described earlier, action updates require a constant number of floating point operations---\num{122} floating point additions and \num{2} multiplications per update of $\wvec{}$, above the \num{140} additions required to choose an action.
The vast majority of these operations can be vectorised trivially, if such hardware is present.
Action computation for SPF is cheaper still, requiring only \num{42} additions per action required.
Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
We believe that it may be possible to implement similar behaviour through application of programmable data planes.

@@

%(rather than na\"{i}ve simulation, blind ML applications etc.) and choose well-considered pathways to solution. \emph{Call-to-action}?

@@

%?? IDEA: try out average reward, TD($\lambda$) methods as future work...
The extant weaknesses invite many further improvements worth investigation.
A natural research direction to enhance this work would be the combination of the classic function approximation techniques we make use of alongside the improved algorithms that the RL community has introduced in the past few years.
Actor-critic methods, or algorithms based on eligibility traces are good candidates for investigation.

@@

%While this work still trails behind the performance of exact DDoS flow detection mechanisms, w
We hope it is clear that the reinforcement learning paradigm holds promise and can inspire further innovation.
We still offer advantages above existing works, such as potentially more flexible deployment and automatically learned low-overhead decision-making, without requiring many of the same network resources or capabilities that they rely upon.
It's hoped that more research in this direction will open the door to works which \emph{respect the complexity of the network environment}; evolving topologies, natural change in traffic and protocol distributions, and the mutation of attacks.

\section*{Acknowledgements}
%Our thanks go to Mircea Iordache, Qianru Zhou, Charles Rutherford, Marco Cook and Richard Cziva for their advice, comments and technical assistance.
Our thanks go to Anonymous, Anonymous, Anonymous, Anonymous and Anonymous for their advice, comments and technical assistance.
%Additional thanks \emph{would} go out to my anonymous reviewers, had I any of them.
This work was supported by a nameless funder [grant number 0].
